Summer 2024 Report

### 2025-01-20 Relations/Report.tex
ARS
Local implications table work

@@

@@

  \item $\forall$ (R : $\mathscr{R}$ A) $\to$ $\forall$ is R -WN x $\to$ is R -UN x $\to$ is R -CR x
  \item $WN_R$ and $UN_R$ $\implies$ $CR_R$

@@

    \end{center}
Counterexample to:
\begin{itemize}
  \item WN $\implies$ SN
  \item WN and CR $\implies$ SN
  \item WN, CR, and bounded $\implies$ SN
  \item WN, UN, and bounded $implies$ SN
\end{itemize}

@@

@@

@@

\subsection{Main themes}
\begin{enumerate}
  \item Clean up the implications around recurrence properties:
  \begin{itemize}
    \item RP iff cofinality and Boundedness?
    \item What is the ``recurrent'' generalization of WN?
    \item What is the ``recurrent'' generalization of SN?
    \item The role of WCR in the proof of our generalization of Thm. 1.2.3(iii)
  \end{itemize}
  \item Complete the ``design'' of the chart(s) with implications?
  \begin{itemize}
    \item   Look at Exercise 1.3.22 (Show UN $\to$ UN$^\to$ but not conversely. Show a counterexample to WCR and UN$^\to$ $\to$ UN. Find a counter example to WCR and WN $\to$ Ind)
  \end{itemize}
  \item Complete the proof of Theorem 1.2.3
  \item Address outstanding questions about wellfoundness implications.
  \begin{itemize}
    \item WFseq implies WFmin? (Assuming decMin)
    \item WFacc implies $WFmin_0$? (Assuming decMin)
    \item DecMin suffices for $SN\to WN$, but what is the optimal \emph{necessary} condition?
    \item Relationship between these, and De Morgan's laws.
  \end{itemize}
  \item Complete the classical properties section
  \item Complete the guide to the code section
  \item Miscellaneous
  \begin{itemize}
    \item   Make the first argument to "being a normal form" implicit?
    \item   Define: Inc in report
    \item   Define: Ind in report
    \item   Define: $\mathrm{CR}^{\le 1}$ in report
    \item   More laws about closure operations: Monotonicity, idempotency
    \item   Tidy up and standardize the notation used in ars.agda and other agda files

  \end{itemize}
\end{enumerate}

>>>>

\begin{dfn}
  A relation $R$ has the \emph{recurrence property} (RP) if
 $a \tto b$ and $a \tto c$  $\implies$
 $c \tto b$, for all $a,c \in A$ and all $b \in A$ with the recurrence property.
\end{dfn}
\sacomment{This mirrors our Weak normal form property but for recurrence. Which would be a suitable name?}

\begin{dfn}
  An element $a \in A$ is \emph{$R$-weakly recurrent} ($WR_{R}$)if $a \tto b$ for some recurrent element $b \in A$.

  A relation $R$ is weakly recurrent (WR) if every $a \in A$ is $R$- weakly recurrent.
\end{dfn}

\begin{dfn}
  An element $a \in A$ is \emph{$R$-strongly recurrent} ($SR_R$) if every reduction sequence starting from $a$ reaches a recurrent element in a finite number of reductions.

  A relation $R$ is strongly recurrent (SR) if every
  $a \in A$ is $R$-strongly recurrent.
\end{dfn}
\sacomment{Not sure on this defn. Is it right to state it requires a finite number of steps}

@@

\subsection{Hierarchy}
The properties of strong normality and confluence are the two properties of greatest interest. As such, there is 
value in determining which other properties combine to imply one these properties. This motivated the development of the 
following hierarchies.

\subsubsection{Confluence hierarchy}

Unique Normal Form Property $<$ Weak Normal Form Property $<$ Recurrence Property $<$ Confluent

\subsubsection{Normalization hierarchy}

Weakly Recurrent $<$ Weakly Normal $<$ Strongly Recurrent $<$ Weakly Normal and Strongly Recurrent $<$ Strongly Normal

The property of being strongly recurrent implies the property of being weakly normal only when there is also the classical property 
of decidability of minimality for a given element.

\begin{table}[h!]
  \centering
  \caption{Local implications}
  \begin{tabular}{|l|c|c|c|c|c|}     
  \hline
  \textbf{CR} & CR, WR  & CR, SR  &   &   &   \\
  \hline
  \textbf{RP} & CR, WR  & CR, SR  &   &   &   \\
  \hline
  \textbf{NP} & NP, WR  & CR, WN  &   &   &   \\
  \hline
  \textbf{UN} & UN, WR  & UN, WN  &   &   &   \\
  \hline
  & \textbf{WR} & \textbf{WN} & \textbf{SR} & \textbf{WN}$\boldsymbol{\land}$\textbf{SR} & \textbf{SN} \\
  \hline  
  \end{tabular}
\end{table}

\begin{table}[h!]
  \centering
  \caption{Local implications - Counterexamples}
  \begin{tabular}{|l|c|c|c|c|c|}     
  \hline
  \textbf{CR} & CR, CE-9  & CR, CE-6  & WIP , CE-9  &   &   \\
  \hline
  \textbf{RP} & CR, CE-9  & CR, CE-6  & WIP , CE-9 &   &   \\
  \hline
  \textbf{NP} & CE-10, CE-9  & CR, CE-6  & CE-11 , CE-9 &   &   \\
  \hline
  \textbf{UN} & CE-3, CE-9  & CE-3, CE-6  & CE-3, CE-9  &   &   \\
  \hline
  & \textbf{WR} & \textbf{WN} & \textbf{SR} & \textbf{WN}$\boldsymbol{\land}$\textbf{SR} & \textbf{SN} \\
  \hline  
  \end{tabular}
\end{table} 
\sacomment{We can potentially merge the two tables. Is it the case that we never move one step up the hierarchy? Is it always a case that if we move one step we move up all the steps, if that is the case then we just need to either express the relevant counterexample, or express the property attained.}

@@

  \item $\forall$ (R : $\mathscr{R}$ A) $\to$ $\forall$ is R -WN x $\to$ is R -UN x $\to$ is R -CR x
  \item $WN_R$ and $UN_R$ $\implies$ $CR_R$
  \item (Local) WR and UN $\implies$ NP 
  \item (Local) WN and UN $\implies$ SR

@@

    \end{center}
Counterexample to:
\begin{itemize}
  \item WN $\implies$ SN
  \item WN and CR $\implies$ SN
  \item WN, CR, and bounded $\implies$ SN
  \item WN, UN, and bounded $implies$ SN
  \item (Local) WN and UN $\implies$ SR
  \item (Local) WN and NP $\implies$ SR
  \item (Local) WN and RP $\implies$ SR
  \item (Local) WN and CR $\implies$ SR
\end{itemize}

@@

\subsection*{Counterexample 9}
\begin{center}

  \begin{tikzcd}[node distance=2cm, auto]
      % Nodes
      \node (a) {a};
      \node (b) [right of=a, xshift=1cm] {b};

      % Arrows
      \draw[->] (a) to [bend left] (b);
      \draw[->] (b) to [bend left] (a);
  \end{tikzcd}

\end{center}
Counterexample to:
\begin{itemize}
  \item (Local) WR and UN $\implies$ WN 
  \item (Local) SR and UN $\implies$ WN 
  \item (Local) SR and NP $\implies$ WN 
  \item (Local) SR and RP $\implies$ WN 
  \item (Local) SR and CR $\implies$ WN 
\end{itemize}

\subsection*{Counterexample 10}
\begin{center}
  \begin{tikzpicture}
      % Nodes
      \node (a) {a};
      \node (b) [left of=a, xshift=-1cm, yshift=-1cm] {b};
      \node (c) [left of=b] {c};
      \node (f0) [right of=a, xshift=1cm, yshift=-1cm] {$f_0$};
      \node (f1) [right of=f0, xshift=1cm] {$f_1$};
      \node (f2) [right of=f1, xshift=1cm] {$f_2$};
      \node (dots) [right of=f2, xshift=1cm] {...};

      % Arrows
      \draw[->] (a) -- (b);
      \draw[->] (b) to [bend left] (c);
      \draw[->] (c) to [bend left] (b);
      \draw[->] (a) -- (f0);
      \draw[->] (f0) -- (f1);
      \draw[->] (f1) -- (f2);
      \draw[->] (f2) -- (dots);
  \end{tikzpicture}
\end{center}

Counterexample to:
\begin{itemize}
  \item (Local) WR and NP $\implies$ RP 
\end{itemize}

\subsection*{Counterexample 11}
\begin{center}
  \begin{tikzpicture}
      % Nodes
      \node (a) {a};
      \node (b) [left of=a, xshift=-1cm, yshift=-1cm] {b};
      \node (c) [left of=b] {c};
      \node (d) [right of=a, xshift=1cm, yshift=-1cm] {d};
      \node (e) [right of=d] {e};

      % Arrows
      \draw[->] (a) -- (b);
      \draw[->] (b) to [bend left] (c);
      \draw[->] (c) to [bend left] (b);
      \draw[->] (a) -- (d);
      \draw[->] (d) to [bend left] (e);
      \draw[->] (e) to [bend left] (d);

  \end{tikzpicture}
\end{center}

Counterexample to:
\begin{itemize}
  \item (Local) SR and NP $\implies$ RP 
\end{itemize}

@@

\subsection{Additional Remarks}

\texttt{isWFmin} is only provably when restricted to \emph{decidable} predicates, and this
restriction is also necessary: in general, WFmin for all P implies all predicates are decidable.

@@

\subsection{Main themes}
\begin{enumerate}
  \item Clean up the implications around recurrence properties:
  \begin{itemize}
    \item RP iff cofinality and Boundedness?
    \item What is the ``recurrent'' generalization of WN?
    \item What is the ``recurrent'' generalization of SN?
    \item The role of WCR in the proof of our generalization of Thm. 1.2.3(iii)
  \end{itemize}
  \item Complete the ``design'' of the chart(s) with implications?
  \begin{itemize}
    \item   Look at Exercise 1.3.22 (Show UN $\to$ UN$^\to$ but not conversely. Show a counterexample to WCR and UN$^\to$ $\to$ UN. Find a counter example to WCR and WN $\to$ Ind)
  \end{itemize}
  \item Complete the proof of Theorem 1.2.3
  \item Address outstanding questions about wellfoundness implications.
  \begin{itemize}
    \item WFseq implies WFmin? (Assuming decMin)
    \item WFacc implies $WFmin_0$? (Assuming decMin)
    \item DecMin suffices for $SN\to WN$, but what is the optimal \emph{necessary} condition?
    \item Relationship between these, and De Morgan's laws.
  \end{itemize}
  \item Complete the classical properties section
  \item Complete the guide to the code section
  \item Miscellaneous
  \begin{itemize}
    \item   Make the first argument to "being a normal form" implicit?
    \item   Define: Inc in report
    \item   Define: Ind in report
    \item   Define: $\mathrm{CR}^{\le 1}$ in report
    \item   More laws about closure operations: Monotonicity, idempotency
    \item   Tidy up and standardize the notation used in ars.agda and other agda files
  \end{itemize}
\end{enumerate}

\subsection{12.23}
\begin{enumerate}
  \item Prove isWFseq implies isWFmin1
\end{enumerate}
